---
title: "Installing and Managing Snippet Modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing and Managing Snippet Modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(snippets)
```

# Overview

The `snippets` package provides a powerful module system that allows you to install, manage, and combine individual snippet collections. This modular approach gives you fine-grained control over which snippets are active in your RStudio environment.

## What are Snippet Modules?

Snippet modules are individual `.snippets` files that contain related code snippets. Instead of managing one large snippet file, you can:

- Install specific modules (e.g., just `dplyr` snippets)
- Combine multiple modules automatically
- Update individual modules without affecting others
- Remove modules you no longer need

## Quick Start

### Install Modules

Install one or more snippet modules:

```{r}
# Install a single module
install_snippet_modules("dplyr", type = "r")

# Install multiple modules at once
install_snippet_modules(c("dplyr", "ggplot2", "tidyr"), type = "r")

# Install markdown modules
install_snippet_modules("markdown", type = "markdown")
```

### List Available Modules

See what modules are available:

```{r}
# List all available R modules
list_snippet_modules(type = "r")

# List all available markdown modules  
list_snippet_modules(type = "markdown")
```

### View Installed Modules

Check which modules are currently installed:

```{r}
# Show active modules for R snippets
show_active_modules(type = "r")

# Show active modules for markdown snippets
show_active_modules(type = "markdown")
```

### Remove Modules

Remove modules you no longer need:

```{r}
# Remove a single module
remove_snippet_modules("tidyr", type = "r")

# Remove multiple modules
remove_snippet_modules(c("dplyr", "ggplot2"), type = "r")
```

## Detailed Usage

### Installation Options

The `install_snippet_modules()` function provides several options:

```{r}
install_snippet_modules(
  modules = c("dplyr", "ggplot2"),  # Modules to install
  type = "r",                      # Snippet type ("r", "markdown", etc.)
  source = "package",              # Source ("package", "local", "github", "url")
  backup = TRUE,                   # Create backup before installation
  force_update = FALSE             # Force reinstall existing modules
)
```

#### Parameters Explained:

- **`modules`**: Character vector of module names to install
- **`type`**: Snippet type - usually "r" or "markdown"
- **`source`**: Where to get modules from (currently supports "package")
- **`backup`**: Whether to create a backup of existing snippets (recommended)
- **`force_update`**: Reinstall modules even if already installed

### Force Updates

If you want to update existing modules:

```{r}
# Force update an existing module
install_snippet_modules("dplyr", type = "r", force_update = TRUE)
```

### Working Without Backups

To skip backup creation (not recommended):

```{r}
install_snippet_modules("ggplot2", type = "r", backup = FALSE)
```

## How Module Installation Works

When you install modules, the system:

1. **Validates** that requested modules exist
2. **Creates a backup** of your current snippets (if enabled)
3. **Copies module files** to a local module directory
4. **Updates the registry** to track installed modules
5. **Recomposes** your main snippet file by combining all installed modules
6. **Notifies you** to restart RStudio for changes to take effect

### Module Registry

The package maintains a registry of installed modules that tracks:

- Module name and type
- Installation date
- Source location
- File paths

This registry ensures consistent module management across sessions.

### File Composition

Your main snippet file (e.g., `r.snippets`) is automatically composed from all installed modules. Each time you install or remove modules, this file is regenerated.

## Available Modules

The package comes with several built-in modules:

### R Modules

- **`dplyr`**: Data manipulation snippets (`filter`, `mutate`, `summarise`, etc.)
- **`ggplot2`**: Plotting snippets (`ggplot`, `geom_*`, `theme_*`, etc.)
- **`plotly`**: Interactive plotting snippets
- **`data-types`**: R data type and structure snippets

### Markdown Modules

- **`markdown`**: Basic markdown syntax snippets
- **`rmd`**: R Markdown specific snippets
- **`qmd`**: Quarto document snippets

You can see the complete list with:

```{r}
# All available modules
list_snippet_modules()
```

## Best Practices

### 1. Always Use Backups

Keep backup creation enabled (the default):

```{r}
# Good - creates backup
install_snippet_modules("dplyr", type = "r", backup = TRUE)
```

### 2. Install Related Modules Together

Install related modules in one call for better organization:

```{r}
# Install tidyverse-related modules together
install_snippet_modules(c("dplyr", "ggplot2", "tidyr", "purrr"), type = "r")
```

### 3. Regular Module Maintenance

Periodically review and clean up your installed modules:

```{r}
# Check what's installed
show_active_modules(type = "r")

# Remove unused modules
remove_snippet_modules("old_module", type = "r")
```

### 4. Restart RStudio

Always restart RStudio after installing or removing modules for changes to take effect.

## Troubleshooting

### Module Not Found

If you get a "Module not found" error:

```{r}
# Check available modules
list_snippet_modules(type = "r")

# Make sure you're using the correct module name
```

### RStudio Not Recognizing Changes

If snippets don't appear after installation:

1. **Restart RStudio completely** (close and reopen)
2. Check that the snippet file was created:

```{r}
# Check if file exists
file.exists(path_rstudio_snippets_file(type = "r"))
```

### Permission Issues

If you encounter permission errors:

- Make sure RStudio is closed during installation
- Run R with appropriate permissions
- Check that the RStudio snippets directory is writable

### Backup Warnings

If you see backup warnings when running outside RStudio, this is normal. The package detects when it's not running in RStudio and adjusts accordingly.

## Integration with Other Functions

Module installation works seamlessly with other package functions:

```{r}
# After installing modules, you can:

# Open the snippet directory
open_rstudio_snippets_dir()

# Edit the main snippet file
open_rstudio_snippets_file(type = "r")

# Manage backups
list_snippet_file_backups(type = "r")
```

## Examples

### Complete Workflow Example

Here's a complete example of module management:

```{r}
# 1. Check what's currently installed
show_active_modules(type = "r")

# 2. See what modules are available
available <- list_snippet_modules(type = "r")
print(available$module)

# 3. Install specific modules
install_snippet_modules(c("dplyr", "ggplot2"), type = "r")

# 4. Verify installation
show_active_modules(type = "r")

# 5. Later, add more modules
install_snippet_modules("plotly", type = "r")

# 6. Remove a module you don't need
remove_snippet_modules("plotly", type = "r")

# 7. Check final state
show_active_modules(type = "r")
```

### Setting Up a New RStudio Installation

For a fresh RStudio setup:

```{r}
# Install essential R modules
install_snippet_modules(
  modules = c("dplyr", "ggplot2", "data-types"),
  type = "r"
)

# Install markdown modules
install_snippet_modules(
  modules = c("markdown", "rmd"),
  type = "markdown"
)

# Verify installation
show_active_modules(type = "r")
show_active_modules(type = "markdown")
```

Remember to restart RStudio after installation for all snippets to become available!
